The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-05-07 RET_Hosts_CWR.tex
Updates from Overleaf

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. Wiseman et al.]{
P. Wiseman$^1$\thanks{E-mail: p.s.wiseman@soton.ac.uk (PW)},
 %M. Pursiainen$^1$,
 %M. J. Childress$^1$,
 %M. Smith$^1$,
 %E. Swann$^2$,
 and Other Authors$^{1,2,3}$,
\newauthor
(The DES Collaboration)
\\
% List of institutions
$^{1}$School of Physics and Astronomy, University of Southampton, Southampton, SO17 1BJ, UK\\
$^{2}$Other Institutions\\
%$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

Of objects passing these cuts, the RETs cluster at shorter timescales and bluer colours than other SNe, even in the observer frame. However, using photometric data points directly has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights, thus making it impossible to measure the peak colour in a number of cases. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a light curve model, hence the rise time values are spread over a wide range. To negate this issue, we use lightcurves interpolated using Gaussian Processes (GP) as presented in \citetalias{Pursiainen2020}. The interpolated light curves have a 0.5 day cadence and every epoch and band has a flux value, and an associated uncertainty. Using this technique, SNe Ia and RETs populate two distinct regions of $g-r$ vs. $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness. We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, and the parameter space can be seen in Fig. \ref{fig:selection}.

@@

\subsubsection{RETs \label{subsubsec:compare_rets}}
Since the DES sample of RETs is by far the largest discovered to date, there is no other large sample of RETs with which to compare host galaxy properties. \citet{Drout2014} present host galaxies of 10 RETs discovered in the Pan-STARRS survey, with measurements of stellar masses, SFRs, and metallicities. We also compare with the host galaxy of SNLS04D4ec \citep{Arcavi2016}. To this we add the low-redshift transient AT2018cow (nicknamed ``The Cow'') with host galaxy photometric measurements from \citet{Perley2019} and metallicity (\citealt{Pettini2004} O3N2) from \citet{Morokuma-Matsui2019}, SN2018gep with data from  \citet{Ho2019}, and ZTF18abvkwla (nicknamed ``The Koala'') from \citet{Ho2020}.

@@

The nearby transient AT2018cow has drawn many comparisons to the cosmological RETs from DES and PS1 \citep[e.g.][]{Perley2019,Margutti2019,Fox2019,Mohan2020} due to its rapid evolution and blue colour. AT2018cow displayed a contracting photosphere as well as evidence for central-engine power alongside an unusual spectrum that showed similarities to broad-lined SNe Ic (SN Ic-bl) at early stages \citep[e.g.][]{Xu2018,Izzo2018} , developing to something entirely different later on \citep{Perley2019} with hints of similarities to interacting SNe Ibn \citep{Fox2019}. There have been several suggestions that AT2018cow is indeed an analogue of the high-z RETs. The host galaxy of AT2018cow appears to be moderately star forming and lies very close to the centre of the SFMS (Figs. \ref{fig:sfms_sfr},\ref{fig:sfms_ssfr}), along with many of the DES RET hosts. However, the host lies significantly above the fiducial MZR in Fig. \ref{fig:mzr}, suggesting that it has an abnormally high metallicity for its stellar mass. This is in contrast to the DES RET hosts, which are systematically less enriched for a given stellar mass.

@@

\section{Conclusions}
\label{sec:conc}
By analysing the host galaxies of 49 rapidly evolving transients (RETs) discovered in the Dark Energy Survey, we have been able to place constraints on the nature of these as-yet unexplained phenomena. We conclude that RETs are strongly linked to massive stars, due to their hosts all exhibiting signatures of star formation. They likely originate from stars more massive, on average, than than those that cause SNe II, and perhaps all SESNe, as they occur in galaxies with higher sSFR. RET hosts are significantly lower in metallicity than SN II hosts, and marginally lower than SESN hosts, suggesting some reliance on rotational energy or other metallicity-dependent effects.
Of the RET analogues discovered in modern large-area, high-cadence surveys, ZTF18abvkwla shares the most similar host galaxy characteristics with the DES RET population. SN2018gep appears in a galaxy too faint to have been detected by DES-SN at the redshift of most of the DES RETs, while the host of AT2018cow is much higher in metallicity.

While current surveys such as ZTF \citep{Bellm2019}, GOTO \citep{Dyer2018}, and BlackGEM \citep{Bloemen2016} are well equipped to find low-redshift RETs, a sample similar to that presented here will likely not be collected until LSST comes online. With several hundreds of objects, detailed studies of RETs and their hosts will be possible in a systematic and more complete manner as has been achieved with LGRBs and SLSNe. At that point, the conclusions of this paper will be able to be tested.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{mc_figs/rets_p16_OH_histfit_final.png}
\caption{Results of the MCMC fits to the PDFs of metallicity for DES RETs and SLSNe, accounting for uncertainties in each bin.
\label{fig:histfit_oh_p16}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[DES Collaboration]{
\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Pursiainen,$^{1}$
M.~Childress,$^{1}$
E.~Swann,$^{2}$
M.~Smith,$^{1}$
L.~Galbany,$^{3}$
C.~Lidman,$^{4}$
T.~M.~Davis,$^{5}$
A.~M\"oller,$^{6}$
B.~P.~Thomas,$^{7,2}$
C.~Frohmaier,$^{2}$
R.~J.~Foley,$^{8}$
S.~R.~Hinton,$^{5}$
L.~Kelsey,$^{1}$
G.~F.~Lewis,$^{9}$
M.~Sako,$^{10}$
D.~Scolnic,$^{11}$
M.~Sullivan,$^{1}$
M.~Vincenzi,$^{2}$
B.~E.~Tucker,$^{4}$
C.~P.~Guti\'errez,$^{1}$
T.~M.~C.~Abbott,$^{12}$
M.~Aguena,$^{13,14}$
S.~Allam,$^{15}$
J.~Annis,$^{15}$
E.~Bertin,$^{16,17}$
S.~Bhargava,$^{18}$
D.~Brooks,$^{19}$
D.~L.~Burke,$^{20,21}$
A.~Carnero~Rosell,$^{7}$
D.~Carollo,$^{7,22}$
M.~Carrasco~Kind,$^{23,24}$
J.~Carretero,$^{25}$
M.~Costanzi,$^{26,27}$
L.~N.~da Costa,$^{14,28}$
H.~T.~Diehl,$^{15}$
P.~Doel,$^{19}$
S.~Everett,$^{8}$
J.~Frieman,$^{15,29}$
J.~Garc\'ia-Bellido,$^{30}$
E.~Gaztanaga,$^{31,32}$
K.~Glazebrook,$^{33}$
D.~Gruen,$^{34,20,21}$
R.~A.~Gruendl,$^{23,24}$
J.~Gschwend,$^{14,28}$
G.~Gutierrez,$^{15}$
K.~Honscheid,$^{35,36}$
D.~J.~James,$^{37}$
K.~Kuehn,$^{38,39}$
N.~Kuropatkin,$^{15}$
M.~Lima,$^{13,14}$
M.~A.~G.~Maia,$^{14,28}$
J.~L.~Marshall,$^{40}$
P.~Martini,$^{35,41}$
F.~Menanteau,$^{23,24}$
R.~Miquel,$^{42,25}$
A.~Palmese,$^{15,29}$
F.~Paz-Chinch\'{o}n,$^{43,24}$
A.~A.~Plazas,$^{44}$
A.~K.~Romer,$^{18}$
E.~Sanchez,$^{45}$
M.~Schubnell,$^{46}$
S.~Serrano,$^{31,32}$
I.~Sevilla-Noarbe,$^{45}$
N.~E.~Sommer,$^{4}$
E.~Suchyta,$^{47}$
M.~E.~C.~Swanson,$^{24}$
G.~Tarle,$^{46}$
D.~L.~Tucker,$^{15}$
T.~N.~Varga,$^{48,49}$
and A.~R.~Walker$^{12}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{Affiliations are listed at the end of the paper}
}
% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

Of objects passing these cuts, the RETs cluster at shorter timescales and bluer colours than other SNe, even in the observer frame. However, using photometric data points directly has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights, thus making it impossible to measure the peak colour in a number of cases. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a light curve model, hence the rise time values are spread over a wide range. To negate this issue, we use lightcurves interpolated using Gaussian Processes (GP) as presented in \citet{Pursiainen2020}. The interpolated light curves have a 0.5 day cadence and every epoch and band has a flux value, and an associated uncertainty. Using this technique, SNe Ia and RETs populate two distinct regions of $g-r$ vs. $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness. We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, and the parameter space can be seen in Fig. \ref{fig:selection}.

@@

\subsubsection{RETs \label{subsubsec:compare_rets}}
Since the DES sample of RETs is by far the largest discovered to date, there is no other large sample of RETs with which to compare host galaxy properties. \citet{Drout2014} present host galaxies of 10 RETs discovered in the Pan-STARRS survey, with measurements of stellar masses, SFRs, and metallicities. We also compare with the host galaxy of SNLS04D4ec \citep{Arcavi2016}. To this we add the low-redshift transient AT2018cow (nicknamed ``The Cow''). The host galaxy of AT2018cow has been studied with photometric measurements \citep{Perley2019} as well as with an integral field spectrograph \citep{Lyman2020}, with consistent results. For our comparison, we use the galaxy-averaged stellar mass and SFR from \citet{Lyman2020} and the metallicity from the host nucleus as reported by both \citet{Morokuma-Matsui2019} and \citet{Lyman2020} as it best represents the method of obtaining spectra for our RET sample. We further compare to SN2018gep with data from  \citet{Ho2019}, and ZTF18abvkwla (nicknamed ``The Koala'') from \citet{Ho2020}.

@@

The nearby transient AT2018cow has drawn many comparisons to the cosmological RETs from DES and PS1 \citep[e.g.][]{Perley2019,Margutti2019,Fox2019,Mohan2020} due to its rapid evolution and blue colour. AT2018cow displayed a contracting photosphere as well as evidence for central-engine power alongside an unusual spectrum that showed similarities to broad-lined SNe Ic (SN Ic-bl) at early stages \citep[e.g.][]{Xu2018,Izzo2018} , developing to something entirely different later on \citep{Perley2019} with hints of similarities to interacting SNe Ibn \citep{Fox2019}. There have been several suggestions that AT2018cow is indeed an analogue of the high-z RETs. The host galaxy of AT2018cow appears to be moderately star forming and lies very close to the centre of the SFMS (\citealt{Lyman2020}; Figs. \ref{fig:sfms_sfr},\ref{fig:sfms_ssfr}), along with many of the DES RET hosts. However, the host lies somewhat above the fiducial MZR in Fig. \ref{fig:mzr}, suggesting that it has an unusually high metallicity for its stellar mass. While consistent with SNe II, this is in contrast to the DES RET hosts which are systematically less enriched for a given stellar mass.

@@

\section{Conclusions}
\label{sec:conc}
By analysing the host galaxies of 49 rapidly evolving transients (RETs) discovered in the Dark Energy Survey, we have been able to place constraints on the nature of these as-yet unexplained phenomena. We conclude that RETs are strongly linked to massive stars, due to their hosts all exhibiting signatures of star formation. They likely originate from stars more massive, on average, than than those that cause SNe II, and perhaps all SESNe, as they occur in galaxies with higher sSFR. RET hosts are significantly lower in metallicity than SN II hosts, and marginally lower than SESN hosts, suggesting some reliance on rotational energy or other metallicity-dependent effects.
Of the RET analogues discovered in modern large-area, high-cadence surveys, ZTF18abvkwla shares the most similar host galaxy characteristics with the DES RET population. SN2018gep appears in a galaxy too faint to have been detected by DES-SN at the redshift of most of the DES RETs, while the host of AT2018cow is higher in metallicity.

While current surveys such as ZTF \citep{Bellm2019}, GOTO \citep{Dyer2018}, and BlackGEM \citep{Bloemen2016} are well equipped to find low-redshift RETs, a sample similar to that presented here will likely not be collected until LSST comes online. With several hundreds of objects, detailed studies of RETs and their hosts will be possible in a systematic and more complete manner as has been achieved with LGRBs and SLSNe.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{mc_figs/rets_p16_OH_histfit_final.png}
\caption{Results of the MCMC fits to the PDFs of metallicity for DES RETs and SLSNe, accounting for uncertainties in each bin.
\label{fig:histfit_oh_p16}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\onecolumn
\input{affiliations.tex}
